Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-10 manuscript/main.tex
empirical covariance as inner product

	Write  $X=(X^{(1)}, \ldots, X^{(K)})$ as a multivariate functional object that belongs to $\mathcal{F} = L^2(\mathcal{T}_1) \times \cdots \times L^2(\mathcal{T}_K)$---a cartesian product of individual $L^2(\mathcal{T}_k)$ spaces. Note that if $K=1$, $X$ reduces to a univariate functional object. We can also express the functional object $X$ evaluated on the multi-dimensional argument $\mathbf{t} = (t_1, \ldots, t_k)^\top \in \mathcal{T} = \mathcal{T}_1 \times \cdots \times \mathcal{T}_K$ as a $K$-dimensional vector $X(\mathbf{t})=(X^{(1)}(t_1), \ldots, X^{(K)}(t_K))^\top$. The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)}) $ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as $\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k$ with norm $\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k)^2 dt_k\}^{1/2}$.

@@

	\begin{equation}
		Y = \sum \limits_{r=1}^p \alpha_r Z_r + \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \beta^{(k)}(t_k) X^{(k)}(t_k) dt_k + \epsilon = \langle \boldsymbol{\alpha}, \mathbf{Z} \rangle +  \langle \beta, X \rangle_{\mathcal{F}}  + \epsilon
		\label{eq: functional model}
	\end{equation}
	where $\boldsymbol{\alpha}=(\alpha_1, \ldots, \alpha_p)^\top \in \mathbb{R}^p$ and $\beta=(\beta^{(1)}, \ldots, \beta^{(K)}) \in \mathcal{F}$ are respectively the regression coefficient vector and function that characterize the effect of scalar and functional predictors on the outcome.

@@

	\begin{definition}[Hilbert space of hybrid objects]\label{def:hilbert_space}
		We define the inner product between any two hybrid objects,
		$\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
		\begin{equation}
			\langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} 
			:= 
			\langle f_1, f_2\rangle_\mathcal{F} 
			+
			\omega \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_{\mathcal{T}_k} f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k
			+
			\omega \sum \limits_{r=1}^p v_{1r}v_{2r}, 
			\label{eq: hybrid inner product}
		\end{equation}
		with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.
	\end{definition}
	In \eqref{eq: hybrid inner product}, 
	$\omega$ is a positive weight that needs to be pre-specified or estimated. It is mainly used to take into account heterogeneity between functional and scalar parts in terms of measurement scale and/or amount of variation (see Section \ref{subsec: Data Preprocessing}). Without loss of generality and for the clarity of illustration, all the following theoretical results will be derived for $\omega=1$. The results remain valid for any positive weights. Then the functional regression model \eqref{eq: functional model} with multiple functional and scalar predictors can be re-expressed in terms of the scalar-on-hybrid regression model as follows
	\begin{equation}
		Y = \langle \boldsymbol{\eta}, \mathbf{W} \rangle_{\mathcal{H}} + \epsilon,
		\label{eq: Hybrid functional model}
	\end{equation}
	where $\boldsymbol{\eta} = (\beta, \boldsymbol{\alpha}) \in \mathcal{H}$ is a hybrid regression coefficient characterizing the association between the hybrid predictor $\mathbf{W}$ and the outcome $Y$.

	The basic idea of partial least squares (PLS) is to simultaneously decompose the hybrid predictor $\mathbf{W}$ and the real outcome $Y$ in terms of zero mean uncorrelated PLS scores $(\rho_l)_{l \in \mathbb{N}}$ with maximum predictive performance as follows
	\begin{equation}
		\mathbf{W}[\mathbf{t}] = \sum \limits_{l=1}^{\infty} \rho_{l} \boldsymbol{\delta}_l[\mathbf{t}] \quad \textnormal{and} \quad Y = \sum \limits_{l=1}^{\infty}\rho_{l} \nu_l + \epsilon,
		\label{eq: decomposition}
	\end{equation}
	where $(\boldsymbol{\delta}_l)_{l \in \mathbb{N}} \in \mathcal{H}$ and $(\nu_l)_{l \in \mathbb{N}} \in \mathbb{R}$ are appropriate bases. The PLS scores are obtained as $\rho_l =  \langle \mathbf{W}^{[l]}, \boldsymbol{\xi}_l \rangle_{\mathcal{H}}$, where $\mathbf{W}^{[l]}$ denotes the residualized predictor sequentially derived as the residual of the regression of $\mathbf{W}^{[l-1]}$ on $\rho_{l-1}$ with $\mathbf{W}^{[1]} = \mathbf{W}$, and $\boldsymbol{\xi}_l=(\psi_l, \boldsymbol{\theta}_l) \in \mathcal{H}$ is the hybrid PLS component, with $\psi_l = (\psi^{(1)}_l, \ldots, \psi_l^{(K)}) \in \mathcal{F}$ and $\boldsymbol{\theta}_l=(\theta_{l1}, \ldots, \theta_{lp})^\top \in \mathbb{R}^p$, sequentially chosen to maximize the squared covariance between $\rho_l$ and the residualized outcome $Y^{[l]}$---i.e., $\textnormal{Cov}^2(\rho_l, Y^{[l]}) = \textnormal{Cov}^2(\langle \mathbf{W}^{[l]}, \boldsymbol{\xi}_l \rangle_\mathcal{H}, Y^{[l]})$. Here, $Y^{[l]}$ is also sequentially obtained as the residual of the regression of $Y^{[l-1]}$ on $\rho_{l-1}$ with $Y^{[1]} = Y$. The sequentially derived hybrid PLS components \{$\boldsymbol{\xi}_l\}_{l=1}^\infty$, are orthonormal in a sense that $\langle \boldsymbol{\xi}_l, \boldsymbol{\xi}_j \rangle_\mathcal{H} = \mathbb{1}(l=j)$.

@@

			\widehat{\boldsymbol{\xi}}_{l}[\mathbf{t}] 

@@

			\langle \mathbf{W}_{1}^{[l]}, \widehat{\boldsymbol{\xi}}_{l} \rangle_{\mathcal{H}},

@@

			\langle \mathbf{W}_{n}^{[l]}, \widehat{\boldsymbol{\xi}}_{l} \rangle_{\mathcal{H}}

@@

			\widehat{\rho}_{il} = \langle \mathbf{W}_{i}^{[l]}, \widehat{\boldsymbol{\xi}}_{l} \rangle_{\mathcal{H}} = \langle X_i^{[l]}, \widehat{\psi}_{l} \rangle_{\mathcal{F}} + \langle \mathbf{Z}_i^{[l]}, \widehat{\boldsymbol{\theta}}_{l} \rangle, 

@@

		where the $l$-th PLS loading $\widehat{\boldsymbol{\xi}}_{l} = (\widehat{\psi}_{l}, \widehat{\boldsymbol{\theta}}_{l}) \in \mathcal{H}$ maximizes $\widehat{\textnormal{Cov}}^2(\langle \mathbf{W}_{i}^{[l]}, \widehat{\boldsymbol{\xi}}_{l} \rangle_{\mathcal{H}}, Y_i^{[l]})=\widehat{\textnormal{Cov}}^2(\widehat{\rho_{il}}, Y_i^{[l]})$. Specifically, we define the $l$-th PLS loadings pointwise as follows:

@@

			\widehat{\boldsymbol{\xi}}_{l}[\mathbf{t}] = \frac{\sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]}[\mathbf{t}]}{\Vert \sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]} \Vert_{\mathcal{H}}} = \left[ \frac{\sum_{i=1}^n Y_i^{[l]} X_i^{[l]}(\mathbf{t})}{\Vert \sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]} \Vert_{\mathcal{H}}}, \frac{\sum_{i=1}^n Y_i^{[l]}\mathbf{Z}_i^{[l]}}{\Vert \sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]} \Vert_{\mathcal{H}}} \right]^\top,

@@

	\subsection{Step 1: Computing smoothed PLS components}  \label{subsec: Step 1: Computing Smoothed PLS components}
	The function part $\psi$ of the hybrid PLS component $\boldsymbol{\xi}$ is inherently infinite-dimensional, rendering its estimation impossible with finite sample size. As such, we first approximate each functional predictor (observed version when $l=1$, or residualized version when $l \ge 2$) as a linear combination of $M$ basis functions and coefficients. 
	Due to the hybrid nature of the predictors, this approximation requires careful explanation, provided in Section \ref{sec: finite basis approximation}. This representation allows us to work with coefficient vectors, simplifying notation. Section~\ref{sec: pls component computation intermediate} outlines the core strategy for computing the PLS components in this reduced space. Building on this, we then introduce our regularized and smoothed estimation procedure, which constitutes the first step of the iterative algorithm.

@@

	Additionally, it serves as a regularization mechanism, borrowing strength across components and helping to prevent overfitting. 	T

@@

	The goal is to find a unit-norm direction $\boldsymbol{\xi} \in \mathcal{H}$, whose functional components lie in $\operatorname{span} \bigl( b_1(t), \ldots, b_M(t) \bigr)$, that maximizes the empirical covariance between the resulting PLS component scores $\langle \widetilde{W}_1, \boldsymbol{\xi} \rangle_{\mathcal{H}}, \ldots, \langle \widetilde{W}_n, \boldsymbol{\xi} \rangle_{\mathcal{H}}$ and the responses $y_1, \ldots, y_n$.
	We denote the solution as
	\begin{equation}
		\hat{\boldsymbol{\xi}} 
		= 
		(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\zeta}}),
	\end{equation}
	where $\hat{\boldsymbol{\zeta}} \in \mathbb{R}^p$ is the scalar part, and each functional part $\hat{\xi}_j \in L^2([0,1])$ lies in $\operatorname{span}(b_1(t), \ldots, b_M(t))$ and is expressed as
	\begin{equation*}
		\hat{\xi}_j(t) := 
		\sum_{l=1}^M d_{jl} \, b_l(t), \quad j = 1, \ldots, K.
	\end{equation*}

	Let $\mathbf{d}_j := (d_{j1}, \ldots, d_{jM})^\top \in \mathbb{R}^M$ and $\mathbf{y} := (y_1, \ldots, y_n)^\top$. 
	Let $B \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions, with entries
	\[
	B_{l, l'} := \langle b_l, b_{l'} \rangle_{L^2([0,1])}, \quad l, l' = 1, \ldots, M.
	\]
	We compute $\mathbf{d}_1, \ldots, \mathbf{d}_K$ and $\hat{\boldsymbol{\zeta}}$ as follows.
	First, we obtain the unnormalized coefficient vectors. For each $j$, the unnormalized vector $\bar{\mathbf{d}}_j$ is computed by solving
	\begin{equation*}
		B \bar{\mathbf{d}}_j = (\Theta_j B)^\top \mathbf{y},
	\end{equation*}
	The unnormalized scalar component is given by
	\begin{equation}
		\bar{\boldsymbol{\zeta}} := Z^\top \mathbf{y}.
	\end{equation}
	Next, we normalize all components with respect to the hybrid Hilbert norm (see Definition~\ref{def:hilbert_space}) to ensure $\hat{\boldsymbol{\xi}}$ has unit length. Specifically, we define
	\begin{equation}\label{def:computed_PLS_direction}
		\mathbf{d}_j := 
		\frac{1}{
			\left(
			\sum_{j=1}^K \bar{\mathbf{d}}_j^\top B \bar{\mathbf{d}}_j +
			\bar{\boldsymbol{\zeta}}^\top \bar{\boldsymbol{\zeta}}
			\right)^{1/2}
		} \bar{\mathbf{d}}_j, \quad j = 1, \ldots, K,
	\end{equation}
	and
	\begin{equation}
		\hat{\boldsymbol{\zeta}} :=
		\frac{1}{
			\left(
			\sum_{j=1}^K \bar{\mathbf{d}}_j^\top B \bar{\mathbf{d}}_j +
			\bar{\boldsymbol{\zeta}}^\top \bar{\boldsymbol{\zeta}}
			\right)^{1/2}
		} \bar{\boldsymbol{\zeta}}.
	\end{equation}

	Details of the hybrid norm computation are provided in Appendix~\ref{section:proof:eigenproblem_derivation}.

	Finally, the estimated $l$th PLS scores for the $n$ observations,
	denoted $\hat{\boldsymbol{\rho}} \in \mathbb{R}^n$
	is computed via hybrid inner product. In a matrix multiplication form, it is computed as:
	\begin{align*}
		\hat{\boldsymbol{\rho}}
		&:=
		\biggl(
		\langle \widetilde{\mathbf{W}}_1, \hat{\boldsymbol{\xi}} \rangle_{\mathcal{H}},
		\ldots,
		\langle \widetilde{\mathbf{W}}_n, \hat{\boldsymbol{\xi}} \rangle_{\mathcal{H}}
		\biggr)^\top
		\\&	=
		\frac{1}{
			\bigl(
			\mathbf{d}^\top \mathbf{d} + 
			\sum_{k=1}^K (\mathbf{d}^{(k)})^\top  \boldsymbol{\Phi}^{(K)}  \mathbf{d}^{(k)} 
			\bigr)^{1/2}
		}
		\biggl(
		\mathbf{Z} \mathbf{d}
		+
		\sum_{k=1}^K \widetilde{\mathbf{C}}^{(k)} \boldsymbol{\Phi}^{(k)} \mathbf{d}^{(k)}
		\biggr)
		. 
	\end{align*}
	The details of this computation are provided in Appendix~\ref{section:proof:eigenproblem_derivation}.
	Although the unnormalized coefficients are computed separately for the functional and scalar part of the PLS component direction, the normalization step couples them, allowing the components to influence each other. This enables the procedure to account for the correlation between functional and scalar components.

	%	consider the basis function expansions of the functional predictors $X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)$, $i=1, \ldots, n$, and those of the functional parts of the PLS component $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k)$, $k=1, \ldots, K$. Let 
	\begin{proposition} \label{Proposition: generalized rayleigh quotient}
		Let $\boldsymbol{\xi} \in \mathcal{H}$ be any unit-norm direction such that its $1$st through $K$th functional components lie in $\operatorname{span} \bigl( b_1(t), \ldots, b_M(t) \bigr)$, respectively.
		Then, the empirical covariance between the projected component $\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}$ and the response $Y$, approximated in terms of the basis, is given by
		\begin{equation*}
			\widehat{\textnormal{Cov}} \left( \langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y \right)
			:=
			\frac{1}{n} 
			\sum_{i=1}^n
			y_i
			\langle \widetilde{\mathbf{W}}_i, \boldsymbol{\xi} \rangle_{\mathcal{H}}.
		\end{equation*}
		Then the PLS component direction 
		$	\hat{\boldsymbol{\xi}}$ 
		computed in \eqref{def:computed_PLS_direction} maximizes $\widehat{\textnormal{Cov}} (\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y )$.
	\end{proposition}

@@

	\subsubsection{Proposed Approach: Regularized and Smoothed Computation} \label{sec: Regularization of PLS Components}
	Although Proposition \ref{Proposition: generalized rayleigh quotient} provides an efficient and principled way to compute the PLS components $\boldsymbol{\xi}$, their functional part $\psi=(\psi^{(1)}, \ldots, \psi^{(K)})$ may not be smooth, which can lead to overfitting and unstable predictions. Thus, in this section, we provide a modification of Proposition  \ref{Proposition: generalized rayleigh quotient} that yields regularized (smoothed) PLS components.

	A widely-used approach to smoothing a function $\psi^{(k)}$ is to penalize its roughness that can be quantified by the integrated squared second derivative:
	\begin{equation}
		\textnormal{PEN} \left(\psi^{(k)} \right) 
		=  \int_{\mathcal{T}_k} \ddot{\psi}^{(k)}(t)^2 dt = \mathbf{a}^{(k)T} \ddot{J}^{(k)} \mathbf{a}^{(k)},
		\label{eq: roughness penalty}
	\end{equation}
	where the last term is obtained using the basis expansion $\psi^{(k)}(t)=\mathbf{a}^{(k)T}\mathbf{b}^{(k)}(t)$.

>>>>

	Write  $X=(X^{(1)}, \ldots, X^{(K)})$ as a multivariate functional object that belongs to $\mathcal{F} = L^2(\mathcal{T}_1) \times \cdots \times L^2([0,1])$---a cartesian product of individual $L^2([0,1])$ spaces. Note that if $K=1$, $X$ reduces to a univariate functional object. We can also express the functional object $X$ evaluated on the multi-dimensional argument $\mathbf{t} = (t_1, \ldots, t_k)^\top \in \mathcal{T} = \mathcal{T}_1 \times \cdots \times [0,1]$ as a $K$-dimensional vector $X(\mathbf{t})=(X^{(1)}(t_1), \ldots, X^{(K)}(t_K))^\top$. The inner product of $f_1=(f_1^{(1)}, \ldots,f_1^{(K)}) $ and $f_2=(f_2^{(1)}, \ldots,f_2^{(K)})$ in $\mathcal{F}$ is defined as $\langle f_1, f_2\rangle_\mathcal{F} =  \sum_{k=1}^K \langle f_1^{(k)}, f_2^{(k)}\rangle_{L^2} = \sum_{k=1}^K \int_0^1 f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k$ with norm $\Vert f_1 \Vert_\mathcal{F} = \langle f_1,f_1 \rangle_\mathcal{F}^{1/2} = \{ \sum_{k=1}^K \int_0^1 f_1^{(k)}(t_k)^2 dt_k\}^{1/2}$.

@@

	\begin{equation}
		Y = \sum \limits_{r=1}^p \alpha_r Z_r + \sum \limits_{k=1}^K \int_0^1 \beta^{(k)}(t_k) X^{(k)}(t_k) dt_k + \epsilon = \langle \boldsymbol{\alpha}, \mathbf{Z} \rangle +  \langle \beta, X \rangle_{\mathcal{F}}  + \epsilon
		\label{eq: functional model}
	\end{equation}
	where $\boldsymbol{\alpha}=(\alpha_1, \ldots, \alpha_p)^\top \in \mathbb{R}^p$ and $\beta=(\beta^{(1)}, \ldots, \beta^{(K)}) \in \mathcal{F}$ are respectively the regression coefficient vector and function that characterize the effect of scalar and functional predictors on the outcome.

@@

	\begin{definition}[Hilbert space of hybrid objects]\label{def:hilbert_space}
		We define the inner product between any two hybrid objects,
		$\mathbf{h}_1 = (f_1, \mathbf{v}_1)$ and $\mathbf{h}_2 = (f_2, \mathbf{v}_2)$, as
		\begin{equation}
			\langle \mathbf{h}_1, \mathbf{h}_2\rangle_{\mathcal{H}} 
			:= 
			\langle f_1, f_2\rangle_\mathcal{F} 
			+
			\omega \langle  \mathbf{v_1}, \mathbf{v_2} \rangle = \sum \limits_{k=1}^K \int_0^1 f_1^{(k)}(t_k) f_2^{(k)}(t_k) dt_k
			+
			\omega \sum \limits_{r=1}^p v_{1r}v_{2r}, 
			\label{eq: hybrid inner product}
		\end{equation}
		with norm $\Vert \cdot \Vert_\mathcal{H} = \langle \cdot,\cdot \rangle_\mathcal{H}^{1/2}$.
	\end{definition}
	In \eqref{eq: hybrid inner product}, 
	$\omega$ is a positive weight that needs to be pre-specified or estimated. It is mainly used to take into account heterogeneity between functional and scalar parts in terms of measurement scale and/or amount of variation (see Section \ref{subsec: Data Preprocessing}). Without loss of generality and for the clarity of illustration, all the following theoretical results will be derived for $\omega=1$. The results remain valid for any positive weights. Then the functional regression model \eqref{eq: functional model} with multiple functional and scalar predictors can be re-expressed in terms of the scalar-on-hybrid regression model as follows
	\begin{equation}
		Y = \langle \boldsymbol{\eta}, \mathbf{W} \rangle_{\mathcal{H}} + \epsilon,
		\label{eq: Hybrid functional model}
	\end{equation}
	where $\boldsymbol{\eta} = (\beta, \boldsymbol{\alpha}) \in \mathcal{H}$ is a hybrid regression coefficient characterizing the association between the hybrid predictor $\mathbf{W}$ and the outcome $Y$.

	The basic idea of partial least squares (PLS) is to simultaneously decompose the hybrid predictor $\mathbf{W}$ and the real outcome $Y$ in terms of zero mean uncorrelated PLS scores $(\rho_l)_{l \in \mathbb{N}}$ with maximum predictive performance as follows
	\begin{equation}
		\mathbf{W}[\mathbf{t}] = \sum \limits_{l=1}^{\infty} \rho_{l} \boldsymbol{\delta}_l[\mathbf{t}] \quad \textnormal{and} \quad Y = \sum \limits_{l=1}^{\infty}\rho_{l} \nu_l + \epsilon,
		\label{eq: decomposition}
	\end{equation}
	where $(\boldsymbol{\delta}_l)_{l \in \mathbb{N}} \in \mathcal{H}$ and $(\nu_l)_{l \in \mathbb{N}} \in \mathbb{R}$ are appropriate bases. The PLS scores are obtained as $\rho_l =  \langle \mathbf{W}^{[l]}, \xi_l \rangle_{\mathcal{H}}$, where $\mathbf{W}^{[l]}$ denotes the residualized predictor sequentially derived as the residual of the regression of $\mathbf{W}^{[l-1]}$ on $\rho_{l-1}$ with $\mathbf{W}^{[1]} = \mathbf{W}$, and $\xi_l=(\psi_l, \boldsymbol{\theta}_l) \in \mathcal{H}$ is the hybrid PLS component, with $\psi_l = (\psi^{(1)}_l, \ldots, \psi_l^{(K)}) \in \mathcal{F}$ and $\boldsymbol{\theta}_l=(\theta_{l1}, \ldots, \theta_{lp})^\top \in \mathbb{R}^p$, sequentially chosen to maximize the squared covariance between $\rho_l$ and the residualized outcome $Y^{[l]}$---i.e., $\textnormal{Cov}^2(\rho_l, Y^{[l]}) = \textnormal{Cov}^2(\langle \mathbf{W}^{[l]}, \xi_l \rangle_\mathcal{H}, Y^{[l]})$. Here, $Y^{[l]}$ is also sequentially obtained as the residual of the regression of $Y^{[l-1]}$ on $\rho_{l-1}$ with $Y^{[1]} = Y$. The sequentially derived hybrid PLS components \{$\xi_l\}_{l=1}^\infty$, are orthonormal in a sense that $\langle \xi_l, \xi_j \rangle_\mathcal{H} = \mathbb{1}(l=j)$.

@@

			\widehat{\xi}_{l}[\mathbf{t}] 

@@

			\langle \mathbf{W}_{1}^{[l]}, \widehat{\xi}_{l} \rangle_{\mathcal{H}},

@@

			\langle \mathbf{W}_{n}^{[l]}, \widehat{\xi}_{l} \rangle_{\mathcal{H}}

@@

			\widehat{\rho}_{il} = \langle \mathbf{W}_{i}^{[l]}, \widehat{\xi}_{l} \rangle_{\mathcal{H}} = \langle X_i^{[l]}, \widehat{\psi}_{l} \rangle_{\mathcal{F}} + \langle \mathbf{Z}_i^{[l]}, \widehat{\boldsymbol{\theta}}_{l} \rangle, 

@@

		where the $l$-th PLS loading $\widehat{\xi}_{l} = (\widehat{\psi}_{l}, \widehat{\boldsymbol{\theta}}_{l}) \in \mathcal{H}$ maximizes $\widehat{\textnormal{Cov}}^2(\langle \mathbf{W}_{i}^{[l]}, \widehat{\xi}_{l} \rangle_{\mathcal{H}}, Y_i^{[l]})=\widehat{\textnormal{Cov}}^2(\widehat{\rho_{il}}, Y_i^{[l]})$. Specifically, we define the $l$-th PLS loadings pointwise as follows:

@@

			\widehat{\xi}_{l}[\mathbf{t}] = \frac{\sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]}[\mathbf{t}]}{\Vert \sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]} \Vert_{\mathcal{H}}} = \left[ \frac{\sum_{i=1}^n Y_i^{[l]} X_i^{[l]}(\mathbf{t})}{\Vert \sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]} \Vert_{\mathcal{H}}}, \frac{\sum_{i=1}^n Y_i^{[l]}\mathbf{Z}_i^{[l]}}{\Vert \sum_{i=1}^n Y_i^{[l]} \mathbf{W}_i^{[l]} \Vert_{\mathcal{H}}} \right]^\top,

@@

	\subsection{Step 1: Computing smoothed PLS components}  \label{subsec: Step 1: Computing Smoothed PLS components}
	The function part $\psi$ of the hybrid PLS component $\xi$ is inherently infinite-dimensional, rendering its estimation impossible with finite sample size. As such, we first approximate each functional predictor (observed version when $l=1$, or residualized version when $l \ge 2$) as a linear combination of $M$ basis functions and coefficients. 
	Due to the hybrid nature of the predictors, this approximation requires careful explanation, provided in Section \ref{sec: finite basis approximation}. This representation allows us to work with coefficient vectors, simplifying notation. Section~\ref{sec: pls component computation intermediate} outlines the core strategy for computing the PLS components in this reduced space. Building on this, we then introduce our regularized and smoothed estimation procedure, which constitutes the first step of the iterative algorithm.

@@

	Additionally, it serves as a regularization mechanism, borrowing strength across components and helping to prevent overfitting.

@@

	The goal is to find a unit-norm direction $\xi \in \mathcal{H}$, whose functional components lie in $\operatorname{span} \bigl( b_1(t), \ldots, b_M(t) \bigr)$, that maximizes the empirical covariance between the resulting PLS component scores $\langle \widetilde{W}_1, \xi \rangle_{\mathcal{H}}, \ldots, \langle \widetilde{W}_n, \xi \rangle_{\mathcal{H}}$ and the responses $y_1, \ldots, y_n$.
	We denote the solution as
	\begin{equation}
		\hat{\xi} 
		= 
		(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\zeta}}),
	\end{equation}
	where $\hat{\boldsymbol{\zeta}} \in \mathbb{R}^p$ is the scalar part, and each functional part $\hat{\xi}_j \in L^2([0,1])$ lies in $\operatorname{span}(b_1(t), \ldots, b_M(t))$ and is expressed as
	\begin{equation*}
		\hat{\xi}_j(t) := 
		\sum_{m=1}^M d_{jm} \, b_m(t), \quad j = 1, \ldots, K.
	\end{equation*}

	Let $\mathbf{d}_j := (d_{j1}, \ldots, d_{jM})^\top \in \mathbb{R}^M$ and $\mathbf{y} := (y_1, \ldots, y_n)^\top$. 
	Let $B \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions, with entries
	\[
	B_{m, m'} := \langle b_m, b_{m'} \rangle, \quad m, m' = 1, \ldots, M.
	\]
	We compute $\mathbf{d}_1, \ldots, \mathbf{d}_K$ and $\hat{\boldsymbol{\zeta}}$ as follows.
	First, we obtain the unnormalized coefficient vectors. For each $j$, the unnormalized vector $\bar{\mathbf{d}}_j$ is computed by solving
	\begin{equation*}
		B \bar{\mathbf{d}}_j = (\Theta_j B)^\top \mathbf{y},
	\end{equation*}
	The unnormalized scalar component is given by
	\begin{equation}
		\bar{\boldsymbol{\zeta}} := Z^\top \mathbf{y}.
	\end{equation}
	Next, we normalize all components with respect to the hybrid Hilbert norm (see Definition~\ref{def:hilbert_space}) to ensure $\hat{\xi}$ has unit length. Specifically, we define
	\begin{equation}\label{def:computed_PLS_direction}
		\mathbf{d}_j := 
		\frac{1}{
			\left(
			\sum_{j=1}^K \bar{\mathbf{d}}_j^\top B \bar{\mathbf{d}}_j +
			\bar{\boldsymbol{\zeta}}^\top \bar{\boldsymbol{\zeta}}
			\right)^{1/2}
		} \bar{\mathbf{d}}_j, \quad j = 1, \ldots, K,
	\end{equation}
	and
	\begin{equation}
		\hat{\boldsymbol{\zeta}} :=
		\frac{1}{
			\left(
			\sum_{j=1}^K \bar{\mathbf{d}}_j^\top B \bar{\mathbf{d}}_j +
			\bar{\boldsymbol{\zeta}}^\top \bar{\boldsymbol{\zeta}}
			\right)^{1/2}
		} \bar{\boldsymbol{\zeta}}.
	\end{equation}
	Details of the hybrid norm computation are provided in Appendix~\ref{section:proof:eigenproblem_derivation}.

The estimated $l$th PLS component score for the $i$th observations,
	denoted $\hat{\rho}_i$
	is computed via hybrid inner product. In a matrix multiplication form, it is computed as:
	\begin{align*}
	\hat{\rho}_i
		&:=
		\langle \widetilde{\mathbf{W}}_i, \hat{\xi} \rangle_{\mathcal{H}}
		=
		\sum_{j=1}^K \Theta_j B  \mathbf{d}_j
		+
			Z \hat{\boldsymbol{\zeta}} 
		. 
	\end{align*}
	The details of this computation are provided in Appendix~\ref{section:proof:eigenproblem_derivation}.
	Although the unnormalized coefficients are computed separately for the functional and scalar part of the PLS component direction, the normalization step couples them, allowing the components to influence each other. This enables the procedure to account for the correlation between functional and scalar components.

	%	consider the basis function expansions of the functional predictors $X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t_k)$, $i=1, \ldots, n$, and those of the functional parts of the PLS component $\psi^{(k)}(t_k) = \sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t_k)$, $k=1, \ldots, K$. Let 
	\begin{proposition} \label{Proposition: generalized rayleigh quotient}
		Let $\xi \in \mathcal{H}$ be any unit-norm direction such that its $1$st through $K$th functional components lie in $\operatorname{span} \bigl( b_1(t), \ldots, b_M(t) \bigr)$, respectively.
		Then, the empirical covariance between the projected component $\langle \widetilde{\mathbf{W}}, \xi \rangle_{\mathcal{H}}$ and the response $Y$, approximated in terms of the basis, is given by
		\begin{equation*}
			\widehat{\textnormal{Cov}} \left( \langle \widetilde{\mathbf{W}}, \xi \rangle_{\mathcal{H}}, Y \right)
			:=
			\frac{1}{n} 
			\sum_{i=1}^n
			y_i
			\langle \widetilde{\mathbf{W}}_i, \xi \rangle_{\mathcal{H}}.
		\end{equation*}
		Then the PLS component direction 
		$	\hat{\xi}$ 
		computed in \eqref{def:computed_PLS_direction} maximizes $\widehat{\textnormal{Cov}} (\langle \widetilde{\mathbf{W}}, \xi \rangle_{\mathcal{H}}, Y )$.
	\end{proposition}

@@

	\subsubsection{Proposed Approach: Regularized and Smoothed Computation} \label{sec: Regularization of PLS Components}
	Although Proposition \ref{Proposition: generalized rayleigh quotient} provides an efficient and principled way to compute the PLS components $\xi$, their functional part $\psi=(\psi^{(1)}, \ldots, \psi^{(K)})$ may not be smooth, which can lead to overfitting and unstable predictions. Thus, in this section, we provide a modification of Proposition  \ref{Proposition: generalized rayleigh quotient} that yields regularized (smoothed) PLS components.

	A widely-used approach to smoothing a function $\psi^{(k)}$ is to penalize its roughness that can be quantified by the integrated squared second derivative:
	\begin{equation}
		\textnormal{PEN} \left(\psi^{(k)} \right) 
		=  \int_0^1 \ddot{\psi}^{(k)}(t)^2 dt = \mathbf{a}^{(k)T} \ddot{J}^{(k)} \mathbf{a}^{(k)},
		\label{eq: roughness penalty}
	\end{equation}
	where the last term is obtained using the basis expansion $\psi^{(k)}(t)=\mathbf{a}^{(k)T}\mathbf{b}^{(k)}(t)$.
